---
title: "Fit models"
tutorial:
    id: "fit-models"
output: 
    learnr::tutorial:
        progressive: true
        allow_skip: false
runtime: shiny_prerendered
drescription: >
    Learn how to create, fit and interact with PM_model objects.
---

## Welcome
Welcome to the Pmetrics model fitting tutorial. In this tutorial, you will learn how to:

* Create and maipulate `PM_data` and `PM_model` objects.
* Use those objects to create a `PM_fit` object.
* Learn how to interact with the `PM_fit` object and execute it.
* Interact with the `PM_result` object.

### Gfortran
The first step to get Pmetrics up and running is to get Gfortran installed in your system. The prefered way to accomplish this is to use Built-in Pmetrics function `PM_build()`: This function will detect your OS and will try to install gfortran on it . If you are using windows and this fails close/re-open Rstudio and try again, windows has some problems to detech environment variables once they change.
If this option fails, here you have other options based on your OS:

* Windows: Install [Chocolatey](https://chocolatey.org/install), and then install [mingw](https://community.chocolatey.org/packages/mingw) using this command: `choco install mingw`. Remember to restart Rstudio after doing this.
* Linux: There are multiple ways to install gfotran on Linux, depending on the specific distribution you're using the name of the package might be different. In the case of Ubuntu and Ubuntu-like distributions gfotran is included in the build-essentials package `sudo apt install build-essentials`.
* Mac: The easiest way to get gfotran installed on a MacOs computer is to use [Homebrew](https://brew.sh/) using this command: `/bin/bash -c "$(curl -fsSL https://raw.githubusercontent.com/Homebrew/install/HEAD/install.sh)"`, and then install gcc using this command: `brew install gcc`.

After installing Gfotran run `PMbuild()`. You should be seeing a message like this one:

```
Compiling 1 of 9: NPprep...
Compiling 2 of 9: NPeng...
Compiling 3 of 9: ITprep...
Compiling 4 of 9: ITeng...
Compiling 5 of 9: ITerr...
Compiling 6 of 9: SIMeng...
Compiling 7 of 9: DOprep...
Compiling 8 of 9: DOeng...
Compiling 9 of 9: mb2csv...
All packages installed and permanent Fortran modules compiled.
```

### R Basics

This is by no means an exhaustive R [tutorial](https://www.w3schools.com/r/), this section will only cover the basics needed to execute pmetrics and to go through the tutorial.

R is a [interpreded](https://en.wikipedia.org/wiki/Scripting_language), dynamic-typed, object oriented programming language, it is mainly used by data scientist an researches all around the world to perform data-analysis, statistical-inference and even artificial intelligence, between other uses.

To write R code you can directly use the console (usually located in the lower left corner in Rstudio) to type the command and then press return to execute it, let's try it! go to Rstudio and execute this little program: `print(paste("2 + 2 equals", 2+2))`.

```{r}
print(paste("2 + 2 equals", 2+2))
```

To write bigger (and to be able to organize them properly) R programmers write their programs in scripts. An script is a plain text file with a .R extension. Those scripts are then read by the R interpreted and get executed line-by-line. The script editor in Rstudio is usually located in the upper right corner of the interface. If it is closed, you can open it by clicking File>New File>R Script.

Let's write a basic script, open the script editor and paste the following content:

```{r}
my_daily_steps <- c(5220,4803,3981,6299,6003,4967,1238)
# Lines that start with a `#` are ignored by the interpreter and are called comments.
# `X <- Y` is the assignment operator, it can be read as `set X to the value of Y`. 
mean_steps <- mean(my_daily_steps)
# In the previous line we are calling a function, a function is a procedure provided by R or by a third-party by using a library.
# You can always access the manual for an specific function by using the `?` operator i.e. `?mean`
# Plotting (`?plot`)
plot(my_daily_steps)
```
You can also use the [pipe operator](https://www.tidyverse.org/blog/2023/04/base-vs-magrittr-pipe/) to execute functions sequentially.
```{r}

# This is equivalent to print(mean(my_daily_steps))
my_daily_steps|>
mean()|>
print()
```

### Data Object

For this tutorial we will be following the `PM_tutorial` function.

Let's start by including the relevant libraries
```{r setup, context="server"}
library(learnr)
library(Pmetrics)
wd<-tempdir()
setwd(wd)
PM_tutorial(auto_confirm=TRUE)
```
The `PM_tutorial` function creates a folder named `Examples` inside your current working directory, because this is an example we are creating that file inside a temporary folder.
Let's look at one of the example data files provided by the Pmetrics tutorial `ex.csv`
```{r}
setwd(wd)
setwd("Examples/src")
read.csv("ex.csv") |>
head()
```
What you are seeing is a csv file with the minimal Pmetrics' data format, further information can be found [here](https://lapkb.github.io/Pmetrics/articles/data.html).

