---
title: "Created Objects"
format:
  html:
    toc: true
    toc-title: "On this page"
    number-sections: true
    code-fold: false
    css: Styles/style.css
bibliography: Citations.bib
vignette: >
  %\VignetteIndexEntry{Created Objects}
  %\VignetteEngine{quarto}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo=F, message=F}
library(Pmetrics)
library(gt)

r_help <- function(pkg, name) {
    glue::glue("[`{name}`](https://rdrr.io/pkg/{pkg}/man/{name}.html)")
}

gh_help <- function(name) {

    glue::glue("[`{name}`](https://lapkb.github.io/Pmetrics/reference/{name}.html)")

}


custom_table <- function(tab){
  #system.file("extData", tab, package = "Pmetrics") %>%
    read.csv(file.path("Data",tab), na.strings = ".") %>% gt() %>%

    tab_style(
      style = list(
        cell_fill(color = "black"),
        cell_text(color = "white", weight = "bold")
      ),
      locations = cells_column_labels(everything())
    )
}
```

## Objects created after a run

After a successful NPAG/NPOD run, an R datafile is saved in the /outputs subdirectory of the newly created numerically ordered folder in the working directory. This file is called "PMout.Rdata". These data files can be loaded by ensuring that the Runs folder is set as the working directory or using the `path` argument to `$fit()`, and then using the Pmetrics command `` `r gh_help("PM_load")` ``.

There are several Pmetrics data objects contained within the Rdata files which are loaded, making these objects available for plotting and other analysis.

`PM_load()` returns a `r gh_help("PM_result")`object. The fields in the object are in the table below.

```{r echo = F, results = 'asis', label= obj_table_fields}
custom_table("PM_result_fields.csv")
```

To access the fields in a `PM_result` object, use the "\$".

```{r echo=T, eval = F, label = obj_access}
res1 <- PM_load(1)
res1$op
res1$op$time
res1$op$data # this accesses the whole observed vs. predicted raw data frame
```

Most of the fields are themselves R6 objects with methods.

```{r echo = T, eval = F, label = obj_methods}
res1 <- PM_load(1)
res1$op
res1$op$summary()
res1$op$plot()
```

The `PM_result` object also contains several methods. The table below lists these methods (functions) and their descriptions.
```{r echo = F, results = 'asis', label= obj_table_methods}
custom_table("PM_result_methods.csv")
```

