---
title: "Created Objects"
output: 
  rmarkdown::html_vignette:
    css: style.css
bibliography: Citations.bib
vignette: >
  %\VignetteIndexEntry{Created Objects}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo = F, message = F}
library(Pmetrics)
```

## Objects created after a run

After a successful IT2B or NPAG run, an R datafile is saved in the
output subdirectory of the newly created numerically ordered folder in
the working directory. After IT2B, this file is called "IT2Bout.Rdata",
and after NPAG it is called "NPAGout.Rdata". These data files can be loaded by ensuring that the
Runs folder is set as the working directory, and then using the Pmetrics
commands `PM_load()` in <span class="r6">R6</span> or `PMload()` for 
<span class="legacy">Legacy</span>.

There are several Pmetrics data objects contained within the Rdata files
which are loaded, making these objects available for
plotting and other analysis. 

<span class="r6">R6</span>

`PM_load()` returns a `PM_result` object. The fields in the object are in the table below.


```{r echo = F, results = 'asis'}
tabObj <- read.csv("Data/ObjectsR6.csv",na.strings=".")
knitr::kable(tabObj)
```

To access the fields in a `PM_result` object, use the "$". 

```{r echo=T, eval = F}
res1 <- PM_load(1)
res1$op
res1$op$time
res1$op$data #this accesses the whole observed vs. predicted data frame
```

Most of the fields are themselves R6 objects with methods.

```{r echo = T, eval = F}
res1 <- PM_load(1)
res1$op
res1$op$summary()
res1$op$plot()
```

<span class="legacy">Legacy</span>

`PMload()` returns a number of objects into the R environment, suffixed by the run number supplied as an argument to `PMload()`. 


```{r echo = F, results = 'asis'}
tabObj <- read.csv("Data/ObjectsLegacy.csv",na.strings=".")
knitr::kable(tabObj)
```


The objects are the same as the fields in a `PM_result` object, except for the *valid* field, which is not a loaded object. The disadvantage of this approach compared to R6 is that there is no indication what is loaded into R with this command. If you had previously added to your run results with `PMsave()`, you would not necessarily know that there were additional results when you executed `PMload()`.


```{r echo = T, eval = F}
PMload(1)
op.1
final.1
cycle.1
cov.1
post.1 #NPAG only
pop.1 #NPAG only
#other results could be included if PMsave() had been run previously
```

Because Legacy Pmetrics works fully under the S3 framework, these objects are passed to functions and do not have methods attached to them.

```{r echo=F, eval = F}
PMload(1)
summary(op.1)
plot(op.1)
#op.1$plot() will not work since op.1 is not an R6 object and has no attached methods
```

 