---
title: "Overview"
output: 
  rmarkdown::html_vignette:
    css: style.css
vignette: >
  %\VignetteIndexEntry{Overview}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo=F, message=F}
library(Pmetrics)
```

## Software engines

There are three main software engines that Pmetrics controls.

* **IT2B** is the ITerative 2-stage Bayesian parametric population PK modeling
program. It is generally used to estimate parameter ranges to pass to
NPAG. It will estimate values for population model parameters under the
assumption that the underlying distributions of those values are normal
or transformed to normal, e.g. log normal.

* **NPAG** is the Non-parametric Adaptive Grid software. It will create a
non-parametric population model consisting of discrete support points,
each with a set of estimates for all parameters in the model plus an
associated probability (weight) of that set of estimates. There can be
at most one point for each subject in the study population. There is no
need for any assumption about the underlying distribution of model
parameter values.

* The **Simulator** is a semi-parametric Monte Carlo simulation software
program that can use the output of IT2B or NPAG to build randomly
generated response profiles (e.g. time-concentration curves) for a given
population model, parameter estimates, and data input. Simulation from a
non-parametric joint density model, i.e. NPAG output, is possible, with
each point serving as the mean of a multivariate normal distribution,
weighted according to the weight of the point. The covariance matrix of
the entire set of support points is divided equally among the points for
the purposes of simulation.

## Pmetrics control functions

<span class="r6">R6</span>

Pmetrics uses`PM_data` to create data objects,  `PM_model` to create model 
objects, and `PM_fit` to create
objects that combine the model with the data, ready to be run (fitted),
generating probability distributions for primary model parameters. 
These are extensively
documented within R by using the `help(command)` or `?command syntax`.

These functions replace the following Legacy functions: `ITrun`, `ERRrun`, 
`NPrun`.

Invoking the simulator in R6 becomes a method attached to `PM_result` objects
or by using `PM_sim$run()` for models, parameter value probability distributions
and template data not derived from a previous fit, e.g. when lifted from an article.


<span class="legacy">Legacy</span>

Pmetrics has groups of R functions named logically to run each of these
programs and to extract the output. Again, these are extensively
documented within R by using the `help(command)` or `?command syntax`.

* `ITrun`, `ITparse`, `ERRrun`

* `NPrun`, `NPparse`

* `PMload`, `PMsave`, `PMreport`

* `SIMrun`, `SIMparse`

### Run functions

<span class="r6">R6</span>

Once a `PM_fit` object is created, which combines a model with a data file,
it can be run by using the syntax `$run()` to access the appropriate function
defined for the `PM_fit` object. 

```{r echo=T, eval=FALSE}
fit1 <- PM_fit$new(model, data)
fit1$run(options)
```


<span class="r6">R6</span> <span class="legacy">Legacy</span>

For IT2B and NPAG, the "run" functions generate batch files, which when
executed, launch the software programs to do the analysis. `$run(engine="err")`
or `ERRrun()` is a special implementation of IT2B designed to estimate 
the assay error polynomial coefficients from the data, when they cannot be calculated from assay validation data (using `makeErrorPoly()`) supplied by the
analytical laboratory. The batch files contain all the information
necessary to complete a run, tidy the output into a date/time stamped
directory with meaningful subdirectories, extract the information,
generate a report, and a saved Rdata file of parsed output which can be
quickly and easily loaded into R. On Mac (Unix) and Linux systems, 
the batch file automatically launches in a Terminal window. 
Prior to v1.9, on Windows systems, the batch file was launched manually, 
but as of v1.9, this manual step is no longer necessary. The execution of
the program to do the actual model parameter estimation is independent
of R, so that the user is free to use R for other purposes.

<span class="r6">R6</span>

For the Simulator, the `$sim` method for a `PM_result` object will execute the 
program directly within R and return a `PM_sim` object. 

```{r echo=T, eval=FALSE}
run1 <- PM_load(1) 
sim1 <- run1$sim(data = "new.csv")
```

The first line loads previous results of run 1 into a PM_result object called run1. The second line uses model and prior in run1 with new data template.

An equivalent method is to run a new `PM_sim` directly with `PM_sim$run()`. It
also returns a `PM_sim` object, i.e. there is no longer any need to run `SIMparse`.

```{r echo=T, eval=FALSE}
sim1 <- PM_sim$run(poppar = list(...), model = "model.txt", data = "new.csv") 
```

See `vignette("simulation")` for details on specifying `poppar` this way.

<span class="legacy">Legacy</span>

To run the simulator in legacy mode, use `SIMrun`. Results are saved to the 
hard drive in the working directory, and can be read with `SIMparse` as
described below.

```{r echo=T, eval=FALSE}
SIMrun(...)
```

### Parse functions

<span class="r6">R6</span> <span class="legacy">Legacy</span>

For all programs, the "parse" functions will extract the primary output
from the program into meaningful R data objects. For IT2B and NPAG in either R6
or legacy modes, this
is done automatically at the end of a successful run, and the objects
are saved in the output subdirectory as *IT2Bout.Rdata* or *NPAGout.Rdata*,
respectively. The parse functions  run automatically and are not
generally necessary for the user to access.

<span class="r6">R6</span> 

For the simulator the results are returned automatically to the object assigned
to contain the output of the simulator, e.g. `sim1` below. 

```{r echo=T, eval=FALSE}
sim1 <- PM_result$sim(...)
```

As mentioned above, there is no longer any
need to use the `SIMparse` legacy function described below, because the 
`SIMrun` and `SIMparse` legacy functions are combined within the `$sim()` method
for `PM_result` objects or the `$run()` method for `PM_sim` objects.

<span class="legacy">Legacy</span>

After running the simulator with `SIMrun`, the results are saved to files 
in the working 
directory. They can be retrieved using `SIMparse`.

```{r echo=T, eval=FALSE}
simdata <- SIMparse("simout1.txt")
```

### Saving functions

<span class="r6">R6</span> 

The Pmetrics R6 objects `PM_fit`, `PM_result`, `PM_sim`, `PM_valid`, and 
`PM_pta` all have a `$save()` method. This method saves the object to the hard drive
in the current working directory by default. The format is .rds which is a 
binary format used by R to save individual objects. The purpose of the `$save()`
method is to enable retrieval of the object at a later time.

<span class="legacy">Legacy</span>

Objects created that are attached to a particular run can be identified by using
the same suffix number as the run. `PMsave` can add these  attached objects to 
the .Rdata file for that run, which is loaded with `PMload`.

```{r echo=T, eval=FALSE}
sim.2 <- SIMparse("simout1.txt")
PMsave(2)
```
In the above example, `sim.2` will be added to the .Rdata file for future
retrieval with `PMload(2)`.

### Loading functions

<span class="r6">R6</span> 

After a successful IT2B or NPAG run, `PM_load` creates a 
`PM_result` object rather than loading run results into the current 
environment and suffixed with the run number as for legacy mode. 

```{r echo=T, eval=FALSE}
res1 <- PM_load(1)
res1$op$plot()
```

`PM_result`, `PM_fit`, `PM_sim`, `PM_valid`, and 
`PM_pta` all have a `$load()` method that takes as its only argument the
full or relative path (if not
in the working directory) and name of an .rds file created by the corresponding
`$save()` method.

```{r echo=T, eval=FALSE}
sim1 <- PM_sim$load("sim.rds")
```


<span class="legacy">Legacy</span> 

For IT2B and NPAG, the `PMload` function can be used to load the .Rdata file
after a successful run. Objects will be loaded
into the current environment in R and suffixed with ".run", where "run" is
the run number.

```{r echo=T, eval=F}
PMload(1)
plot(op.1)
```

For simulations, `SIMparse`, described above, combines the parsing and loading
functions.

```{r echo=T, eval=FALSE}
sim.2 <- SIMparse("simout1.txt")
```

### Report generation

<span class="r6">R6</span> <span class="legacy">Legacy</span>

The `PMreport` function is automatically run at the end of a successful
NPAG and IT2B run, and it will generate an HTML page with summaries of
the run, as well as the .Rdata files and other objects. The default
browser will be automatically launched for viewing of the HTML report
page. <!--It will also generate a .tex file suitable for processing by a
LATEX engine to generate a pdf report.--> See the [Pmetrics
Outputs](#outputs) section.

## Other functions

<span class="r6">R6</span> <span class="legacy">Legacy</span>

Within Pmetrics there are also functions to manipulate data .csv files
and process and plot extracted data.

### Data manipulation
```{r echo=F}
data_comp <- read.csv("Data/RLcomp_data.csv",na.strings=".")
knitr::kable(data_comp)
```

### Model selection and diagnostics
```{r echo=F}
mod_comp <- read.csv("Data/RLcomp_valid.csv",na.strings=".")
knitr::kable(mod_comp)
```

### Other functions
```{r echo=F}
oth_comp <- read.csv("Data/RLcomp_other.csv",na.strings=".")
knitr::kable(oth_comp)
```


<!-- * Process data: `makeAUC`, `makeCov`, `makeCycle`, `makeFinal`, `makeOP`, `makePTA`, -->
<!-- `makeErrorPoly` -->

<!-- * Plot data: `plot.PMcov`, `plot.PMcycle`, `plot.PMfinal`, `plot.PMmatrix`, -->
<!-- `plot.PMop`,`plot.PMsim`, `plot.PMvalid`, `plot.PMpta` -->


<!-- * Pmetrics function defaults: `setPMoptions`, `getPMoptions` -->

Again, all functions have extensive help files and examples which can be
examined in R by using the `help(command)` or `?command` syntax.