---
title: "Learn Pmetrics"
output: learnr::tutorial
runtime: shiny_prerendered
description: "A tutorial for learning Pmetrics, a package for pharmacometric modeling and simulation in R."
---

```{r setup, include=FALSE}
library(learnr)

knitr::opts_chunk$set(echo = FALSE)
```


## Introduction

This learnr tutorial will introduce you to Pmetrics. We asssume you know
your way around the Rstudio IDE, and have some (very) basic knowledge of R. For 
example, you should understand the difference between R scripting and the R
console. 

As a reminder, in R code, lines that start with "#" are comments and ignored by R.  
Methods to execute R after placing your cursor anywhere in a line of R
code are as follows:

* Windows
  - The Run button above right
  - Ctrl-Enter
* Mac
  - The Run button at the top
  - Command-Return
      
This script also serves to introduce several R programming functions
and techniques. For any function, you can get help by typing ?function_name
in the R console (the lower left window pane in RStudio).

## Loading libraries

At the start of every session, you need to load Pmetrics into memory.  
You must include this code at the beginning of every script. Another incredibly
useful set of packages is loaded with the meta-package, `tidyverse`, which
contains many useful functions for data manipulation and visualization. Pmetrics
uses tidyverse conventions throughout.

```{r}
library(Pmetrics)
library(tidyverse)
```

Let's walk through several exercises to showcase Pmetrics functionality.  The exercises
are designed to be run in the tutorial, so you can see the output as you go.  You can
also run the exercises in a separate R script or by copying the code into the console.

## Exercises

### Exercise 1 - NPAG run 


```{r}
# EXAMPLE NPAG RUN - tlag, ka, kel, vol
```

It is useful to annotate your runs with comments in your script as above, 
so that you can remember what you did later!

Tell R where your working directory is going to be.
Windows users:  Make sure that you separate directories with a
forward slash "/" or double backslashes "\\". Unfortunately, Windows is the only OS that uses
backslashes "\", so R conforms to Unix/Linux style.

```{r}
wd <- "##WD##"
```




## Topic 1

### Exercise 

*Here's a simple exercise with an empty code chunk provided for entering the answer.*

Write the R code required to add two plus two:

```{r two-plus-two, exercise=TRUE}

```

### Exercise with Code

*Here's an exercise with some prepopulated code as well as `exercise.lines = 5` to provide a bit more initial room to work.*

Now write a function that adds any two numbers and then call it:

```{r add-function, exercise=TRUE, exercise.lines = 5}
add <- function() {
  
}
```

## Topic 2

### Exercise with Hint

*Here's an exercise where the chunk is pre-evaluated via the `exercise.eval` option (so the user can see the default output we'd like them to customize). We also add a "hint" to the correct solution via the chunk immediate below labeled `print-limit-hint`.*

Modify the following code to limit the number of rows printed to 5:

```{r print-limit, exercise=TRUE, exercise.eval=TRUE}
mtcars
```

```{r print-limit-hint}
head(mtcars)
```

### Quiz

*You can include any number of single or multiple choice questions as a quiz. Use the `question` function to define a question and the `quiz` function for grouping multiple questions together.*

Some questions to verify that you understand the purposes of various base and recommended R packages:

```{r quiz}
quiz(
  question("Which package contains functions for installing other R packages?",
    answer("base"),
    answer("tools"),
    answer("utils", correct = TRUE),
    answer("codetools")
  ),
  question("Which of the R packages listed below are used to create plots?",
    answer("lattice", correct = TRUE),
    answer("tools"),
    answer("stats"),
    answer("grid", correct = TRUE)
  )
)
```

