---
title: "Introduction"
---

```{r}
#| label: setup
#| echo: false
#| message: false
#| eval: true

pmetrics <- function(){
    knitr::asis_output("[Pmetrics]{style=\"color: #841010; font-family: 'Arial', Arial, sans-serif; font-weight: 900;\"}")
}
```


These exercises will help you learn how to use `r pmetrics()` for R. They are designed to be read in sequence, so start with the first one.

When you work on a project, best practice is to use the following workflow:

1.  Create a new directory for your project
2.  Create a new R script or Quarto document in that directory
3.  Build your project code in that script or document.
4.  You can store source data files in a subdirectory, e.g. "data" or "src" and the various runs when you fit a model to the data in another subdirectory, e.g. "Runs".

If you have loaded `r pmetrics()` already with `library(Pmetrics)`, the command `PMtree()` will create a project folder with subfolders for you. To learn more about this function or any other in R, you can get help by typing `?function_name` in the console.

```{r}
#| eval: false
?PMtree
```

We assume you have some familiarity with R and possibly Quarto. If not, there are many good tutorials online.

## How to use these pages

We have designed these pages with the idea that you can copy and paste the code chunks into your own scripts or Quarto documents to see how they work. You can hover over the code chunks to see a copy icon. Click that to copy the code to your clipboard so you can paste it into your own script or document. Alternatively, you can click the "Copy all code" button at the top right of every page to copy all the code chunks on that page to your clipboard.

**Note:** In many cases, we do not evaluate the code chunks because the ouput is too long, so you will need to run them in your own script or document.

```{r}
# Hover over the code chunk to see the copy icon.
```

## Scripting

You can use either an R script or a Quarto document to paste code from these pages or to write your own code. R scripts have the advantage of being simple and easy to use. Quarto documents have the advantage of being able to combine code with formatted text, images, links, etc. You can use either method to work with `r pmetrics()`. Choose the one that works best for you.

### R script

-   Create a new script with *File -\> New File -\> R Script* (Rstudio) or *R File* (Positron).
-   Save the script in your project directory with a name like "Learn.R".
-   You can then copy and paste code chunks from these pages into your script and run them line by line or all at once.
-   It is useful to annotate your code with comments like the one above so you can remember what you did later.

Here's a toy example of an R script.
```{r}
# This is a comment in R. It starts with a # symbol.
# Anything other than a comment should be R code.
a <- 1 + 1
print(a) # This will print the value of a to the console.
```

### Quarto document

-   Create a new document with *File -\> New File -\> Quarto Document* (Rstudio/Positron).
-   Use markdown to add headings, links, images, etc.
-   Insert R chunks to paste copied code or write your own.
-   Execute code from the chunks.
-   Render the document to create a nicely formatted output in HTML, PDF, or Word format.
-   See [Quarto documentation](https://quarto.org/docs/guide) for more information.

Here's a toy example of a Quarto document. It mixes text formatted in markdown and R code chunks.

```{r}
#| echo: false
#| results: asis
#| eval: true
cat("````qmd\n",
"---\n",
"title: \"My Project\"\n",
"format: html\n",
"---\n\n",
"## Introduction\n",
"This is my first Quarto document. Below is some R code.\n\n",
"```{r}\n",
"# this is an R code chunk\n",
"b <- 2 + 2\n",
"print(b)\n",
"```\n\n",
"You can see that the value of b is printed above.\n\n",
"````",
sep = ""
)
```

When rendered, the above will look like this:

## My Project {.unnumbered .unlisted}
### Introduction {.unnumbered .unlisted}
This is my first Quarto document. Below is some R code.
```{r}
#| eval: true
# this is an R code chunk
b <- 2 + 2
print(b)
```
You can see that the value of b is printed above.



