---
title: "Learn Pmetrics"
output: 
  learnr::tutorial:
    progressive: true
runtime: shiny_prerendered
description: "A tutorial for learning Pmetrics, a package for pharmacometric modeling and simulation in R."
---

```{r setup, include=FALSE}
library(learnr)
library(Pmetrics)
library(tidyverse)

knitr::opts_chunk$set(echo = FALSE)
```


## Introduction

This learnr tutorial will introduce you to Pmetrics. We asssume you know
your way around the Rstudio IDE, and have some (very) basic knowledge of R. For 
example, you should understand the difference between R scripting and the R
console. 

As a reminder, in R code, lines that start with "#" are comments and ignored by R.  
Methods to execute R after placing your cursor anywhere in a line of R
code are as follows:

* Windows
  - The Run button above right
  - Ctrl-Enter
* Mac
  - The Run button at the top
  - Command-Return
      
This script also serves to introduce several R programming functions
and techniques. For any function, you can get help by typing ?function_name
in the R console (the lower left window pane in RStudio).

## Loading libraries

At the start of every session, you need to load Pmetrics into memory.  

You must include this code at the beginning of every script. Another incredibly
useful set of packages is loaded with the meta-package, `tidyverse`, which
contains many useful functions for data manipulation and visualization. Pmetrics
uses tidyverse conventions throughout. You will see some package start up messages,
which can be ignored now.

```{r echo = TRUE, eval = FALSE}
library(Pmetrics)
library(tidyverse)
```

Let's walk through several exercises to showcase Pmetrics functionality.  The exercises
are designed to be run in the tutorial, so you can see the output as you go.  You can
also run the exercises in a separate R script or by copying the code into the console.

## Data objects

### Setting the working directory

Tell R where your working directory is going to be by setting the working directory
using the `setwd()` function.  This is the directory where your data and model files
are located.  You can also use the `getwd()` function to see what your current working
directory is.  If you are using RStudio, you can also set the working directory
using the menu: Session -> Set Working Directory -> Choose Directory.

**Note**: Windows users:  Make sure that you separate directories with a
forward slash "/" or double backslashes "\\". Unfortunately, Windows is the only OS that uses
backslashes "\", so R conforms to Unix/Linux style.

Here we show you how to set the working directory to a dummy folder, but in this exercise,
we will use relative paths to necessary files.

```{r echo = TRUE, eval = FALSE}
setwd("data/my_folder")
# Windows: above will work or use setwd("data\\my_folder")
```

### Creating a data object
Pmetrics always needs data and a model to run. Usually you have the data first
and then make the model once you know what the data look like.  The data
object is created with the `PM_data$new()` function, which takes a file name
as an argument.  The file name can be a relative path (relative to the working directory)
or an absolute path.  The file name should be a `.csv` file, or comma-separated values,
which is a common format for data files. It's possible to use `.ssv` (semicolon-separated)
or other formats, and this can be specified by setting global Pmetrics options.
Type `?setPMoptions` in the R console for more information.

Let's create our first data object.

```{r list-files, exercise = TRUE}
# list the files inside the directory
list.files("data/Examples/src")
```

Make a `PM_data` object by reading the "ex.csv" file.

```{r first-data, exercise = TRUE}
# create a new data object by reading a file
exData <- PM_data$new(data = "data/Examples/src/ex.csv")
```

When you create a `PM_data` object, it reads the data file and performs error checks,
trying to fix any that it can. Unless you use `PM_data$new(..., quiet = TRUE)`, Pmetrics
will give you a brief report of any errors it found and fixed, or if it could not fix them.
Please note, that it is possible your data might contain a kind of error that Pmetrics
hasn't been programmed to detect, as it only checks for mistakes made and reported over
the years by users. If you find an error that Pmetrics does not detect, please
report it via an issue in [Github](https://github.com/LAPKB/Pmetrics/issues).

### Examining data objects

`exData` is an R6 object, which means that contains both data and methods to
process that data, for example:

```{r orig-data, exercise = TRUE, exercise.setup = "first-data"}
exData$data # contains your original datafile
```

Compare the standardized data below to your original data above.

```{r standard-data, exercise = TRUE, exercise.setup = "first-data"}
exData$standard_data # contains the standardized and validated data,
```

You can summarize your data.

```{r summary-data, exercise = TRUE, exercise.setup = "first-data"}

exData$summary() # prints the summary of the data to the terminal, or

# another way to do that is using the more common S3 framework in R:
summary(exData)

# To look at the contents of an object:
names(exData)

```

The original data in the `PM_data` object displays in the viewer. The standardized data
displays by default in the console as a data frame. These defaults can be overridden
by changing arguments to the `$print()` method for `PM_data` objects.

The viewer option doesn't display well in learnr tutorials, so we'll just show the
code but not the output.

```{r echo = TRUE, eval = FALSE}
exData # view the original data in the viewer
exData$print(standard = FALSE, viewer = TRUE) # same as above, arguments set to their defaults
exData$print(standard = TRUE) # view the standardized data in the viewer
```

Here are examples that display fine in learnr tutorials.

```{r data-examples, exercise = TRUE, exercise.setup = "first-data"}
exData$print(viewer = FALSE) # view original data in console
exData$plot() #plot the raw data; more on that later

```


## Model objects

You can specify a model by reading a file or directly as an object. We'll do both.
The following code creates the same model we will read later from a file.
The advantage of creating them in R is that one does not need to copy model
files into folders to provide necessary inputs.

### Models specified in R with code

#### A note about R functions and arguments

First, we'll look at specifying a model as R code directly. However, 
make sure you understand how function arguments work
in R. If you supply argument values in the same order as defined in the function,
without skipping any, R allows you to omit the names of the arguments. 

On the other hand, if you skip any arguments you must specify the names of subsequent arguments so R doesn't get confused. Similarly, if you supply argument values out of order,
you must also include the names of any misordered arguments.

#### Argument details for a model in Pmetrics

The argument order for the function to make a model in Pmetrics is:

* file The name of a text file with a model. This can be the only argument, e.g.
`PM_model$new("model.txt")`
* pri A vector of primary parameters, which are the parameters whose distributions will be estimated.
* cov A vector of covariates, which are in the same order and spelling as in the data.
* sec A function that defines any secondary parameters, which are derived from primary parameters and covariates. Secondary parameters are not estimated, but are used in the model,
such as defining volume of distribution as a function of weight.
* tem A character vector primarily used to select one of the predefined templates
that contain equations for algebraic models that do not use differential equations and 
therefore run much faster. The model templates Pmetrics
knows about can be accessed with `model_lib()`. Models defined only by differential
equations should have "ode" in the `tem` item. If you omit the `tem` argument,
Pmetrics will add it according to the following rules:
  * If the model has no differential equations, the names of the `pri` and `sec`
  parameters will be used to determine the appropriate algebraic template name.
  Specifically, the names Pmetrics is looking for are in the *Parameters* column
  of the table returned by `model_lib()`.
  * If the model has differential equations, the argument will be `tem = "ode"`.
  * If there are no differential equations, and Pmetrics cannot find the appropriately
  named parameters, the argument will be `tem = "diirect"`, which means Pmetrics
  is expecting a direct solution in the `out` argument. This is quite rare.
* eqn A function containing  differential equations that define the model. 
The equations are written in R syntax, so you can use any R function. 
If you select an algebraic model in the `tem` field, you do not need to include 
equations here, but Pmetrics will add them to make the structure
of the model more clear, both by the equations themselves and by using them to 
plot a diagram of the model via the `$plot()` method for `PM_model` objects. 
* lag A function that defines which parameter(s) define the lag time or delay
in absorption for drug(s) in your model. The function is written in R syntax.
* fa A function that defines which parameter(s) define the bioavailability or
fraction absorbed for drug(s) in your model. The function is written in R syntax.
* ini A function that defines which parameter(s) define the initial conditions for
one or more compartments in your model. The function is written in R syntax.
* out A function defining the model output equations. The function is written in R syntax.
* err A vector defining the error model for the output equations. The function is written in R syntax.

The next section provides an example.

#### R code for a model in Pmetrics

Running the code below will compile the model in Rust, print it in the console,
and then plot it so you can see the structure.

```{r mod, exercise = TRUE, eval = FALSE}
mod1 <- PM_model$new( # create a new model object
  
  # primary parameters whose distributions will be estimated
  # here, the prior distributions are specified using the ab() and msd() functions
  
  pri = c( 
    Ka = ab(0.1, 0.9), # the ab() function specifies a possible range for the parameter
    Ke = ab(0.001, 0.1),
    V0 = ab(30, 120),
    lag1 = msd(2, 0.66) # the msd() function specifies a mean, standard deviation
  ), # note the comma which separates the arguments
  
  # list any covariates in the same order as in your data
  
  cov = c(
    "WT", # default without "!" means values will be linearly interpolated between entries
    "AFRICA!", # the "!" indicates this covariate should be held constant until the next entry
    "AGE",
    "GENDER",
    "HEIGHT"
  ),
  sec = function(){
    V = V0 * WT/70 # this is written in R syntax
  }
  
  # see model_lib() for available templates
  tem = "one_comp_iv", 
  
  # the equations that define the model, written in R syntax
  # if you include a template in tem, these will be added for you
  eqn = function(){
    dX[1] = -Ka*X[1] # dX[1] means dX/dt for compartment 1
    dX[2] = Ka*X[1] - Ke*X[2] # X[1] means amount of drug in compartment 1
  },
  
  lag = function(){
    Tlag[1] = lag1 # set lag time for drug 1 (Tlag[1]) to the pri parameter lag1
  },
  
  
  out = function(){
    Y[1] = X[2]/V # there should be one equation for each output
  },
  
  # a vector of error functions, one per output equation
  err = function(){
    E[1] = proportional(initial = 5, fit = TRUE, assay = c(0.02, 0.05, -0.0002, 0))
  }

)
# look at it
mod1

#plot it
mod1$plot()
```

The function to create a model has several arguments, which are described in the
help, accessed by typing `?PM_model` in the R console. 

Note the last command, which takes advantage of the R6 framework, which is object oriented.
That means that every object in Pmetrics contains data and the methods to act on that data. In this case a `PM_model` object contains a `$plot()` method that displays the model structure in a diagram.

### Models specified in a file

```{r model-text, exercise = TRUE}
# in the working directory we have another file "model.txt" that contains the old
# representation of the same model we previously presented, let's take a look at it.
system("cat data/Examples/src/model.txt")

# PM_model$new() also accepts the path to a model file
# create the same model using this file
mod1b <- PM_model$new("data/Examples/src/model.txt")
mod1b

```

