---
title: "Pmetrics - Report"
author: "Laboratory of Applied Pharmacokinetics and Bioinformatics (LAPKB)"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
params:
  res: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

library(tidyverse)
library(patchwork)
library(DT)
library(Pmetrics)
library(htmltools)
library(crayon)
```


```{r setup-extra}
#res = params$res
res <- Pmetrics::NPex
```

## Summary

```{r}
did_converge <- ifelse(res$NPdata$converge == 1, "converged", "did not converge")
```


The data had `r res$NPdata$nsub` subjects and a total of `r  sum(res$NPdata$nobs, na.rm = TRUE)` observations.

After `r res$NPdata$icyctot`/`r res$NPdata$icycmax` cycles, the model `r did_converge`.

## Observed-predicted plot {.tabset .tabset-pills}

Points should appear close to, and randomly distributed around, the (dashed) identity-line.

``` {r include = FALSE}

template <- c(
  "### Outeq {{outeq}} {.tabset .tabset-pills}\n",
  "#### Population {.tabset .tabset-pills}\n",
  "##### Median\n",
  "```{r message = F}\n",
  "res$op$plot(icen = 'median', pred.type = 'pop', outeq = {{outeq}},\n
  stats = list(x = 1, y = 0.1, font = list(size = 9)))\n",
  "```\n",
  "##### Mean\n",
  "```{r message = F}\n",
  "res$op$plot(icen = 'mean', pred.type = 'pop', outeq = {{outeq}},\n
  stats = list(x = 1, y = 0.1, font = list(size = 9)))\n",
  "```\n",
  "#### Posterior {.tabset .tabset-pills}\n",
  "##### Median\n",
  "```{r message = F}\n",
  "res$op$plot(icen = 'median', pred.type = 'post', outeq = {{outeq}},\n
  stats = list(x = 1, y = 0.1, font = list(size = 9)))\n",
  "```\n",
  "##### Mean\n",
  "```{r message = F}\n",
  "res$op$plot(icen = 'mean', pred.type = 'post', outeq = {{outeq}},\n
  stats = list(x = 1, y = 0.1, font = list(size = 9)))\n",
  "```\n",
  "\n"
)

numeqt <- max(res$op$data$outeq)
p <- lapply(1:numeqt, function(i) {
  knitr::knit_expand(
    outeq = i,
    text = template)
})

```
`r knitr::knit_child(text = unlist(p))`

## Prediction metrics {.tabset .tabset-pills}

*Measures of bias. Others are measures of imprecision.

^Same units as observations and predictions. Others are squared or dimensionless.
```{r}
#summary function
metrics <- function(icen, outeq){
  pe <- res$op$summary(icen = icen, outeq = outeq)$pe %>%
    rename(`Mean PE (prediction - observation)*^` = mpe,
           `Mean weighted PE*^` = mwpe,
           `Mean squared PE` = mspe,
           `Mean weighted squared PE` = mwspe,
           `Root mean squared PE^` = rmse,
           `Root mean squared PE (%)` = percent_rmse,
           `Bias-adjusted mean squared PE` = bamspe,
           `Bias-adjusted mean weighted squared PE` = bamwspe
    ) %>% 
    mutate(across(.cols = everything(),
                  .fns = function(x) {
                    round(x, digits = 3)
                  })) %>% 
    t() %>% 
    datatable(
      colnames = c("Metric", "Value"),
      options = list(
        searching = FALSE,
        pageLength = 10,
        lengthChange = FALSE,
        info = FALSE,
        dom = "t",
        autoHideNavigation = TRUE
      )
    )  
  return(pe)
}


```

```{r, results = "asis"}

for(i in 1:res$NPdata$numeqt){
  cat("### Outeq ",i,"{.tabset .tabset-pills}\n")
  cat("\n")
  cat("#### Median\n")
  cat("\n")
  print(tagList(metrics("median", i)))
  cat("\n")
  cat("#### Mean\n")
  cat("\n")
  print(tagList(metrics("mean", i)))
  cat("\n")
}

```

## Prediction residuals {.tabset .tabset-pills}

Residuals should be randomly distributed around zero, with no apparent trend.
For how to interpret these plots, please review our [website](https://lapkb.github.io/Pmetrics). Weighted prediction errors are weighted by the observation standard deviation, i.e. the error polynomial.

``` {r include = FALSE}

template <- c(
  "### Outeq {{outeq}} {.tabset .tabset-pills}\n",
  "#### Population {.tabset .tabset-pills}\n",
  "##### Median\n",
  "```{r message = F}\n",
  "res$op$plot(resid = T, icen = 'median', pred.type = 'pop', outeq = {{outeq}})\n",
  "```\n",
  "##### Mean\n",
  "```{r message = F}\n",
  "res$op$plot(resid = T, icen = 'mean', pred.type = 'pop', outeq = {{outeq}})\n",
  "```\n",
  "#### Posterior {.tabset .tabset-pills}\n",
  "##### Median\n",
  "```{r message = F}\n",
  "res$op$plot(resid = T, icen = 'median', pred.type = 'post', outeq = {{outeq}})\n",
  "```\n",
  "##### Mean\n",
  "```{r message = F}\n",
  "res$op$plot(resid = T, icen = 'mean', pred.type = 'post', outeq = {{outeq}})\n",
  "```\n",
  "\n"
)

numeqt <- max(res$op$data$outeq)
p <- lapply(1:numeqt, function(i) {
  knitr::knit_expand(
    outeq = i,
    text = template)
})

```
`r knitr::knit_child(text = unlist(p), options = list(fig.width = 10))`

## Parameter values and distributions

### Parameter summmaries
The model included the following primary parameters
```{r}
par_lims <- data.frame(
  name = names(res$final$popMean),
  lower = res$final$ab[,1],
  upper = res$final$ab[,2],
  ave = t(res$final$popMean),
  med = t(res$final$popMedian),
  sd = t(res$final$popSD),
  cv = t(res$final$popCV),
  shrink = 100*t(res$final$shrinkage)
) %>%
  mutate(across(4:8,round,3))

par_lims %>% 
  rename(`Parameter` = name,
         `Lower limit` = lower,
         `Upper limit` = upper,
         `Mean` = ave,
         `Median` = X2, #no idea why this gets named X2
         `SD` = sd,
         `CV%` = cv,
         `%shrink` = shrink) %>% 
  datatable(
    options = list(
      searching = FALSE,
      pageLength = 10,
      lengthChange = FALSE,
      info = FALSE,
      dom = "t",
      autoHideNavigation = TRUE
    ),
    rownames = FALSE
  )
```

### Marginal plots
Support points, dashed lines are limits

<!-- ```{r} -->
<!-- res$final$popPoints %>%  -->
<!--   pivot_longer(cols = -prob) %>%  -->
<!--   left_join(par_lims, by = "name") %>%  -->
<!--   ggplot(aes(x = value, y = prob)) + -->
<!--   geom_segment(aes(xend = value, yend = 0)) + -->
<!--   geom_vline(aes(xintercept = lower), lty = "dashed", alpha = 0.5) + -->
<!--   geom_vline(aes(xintercept = upper), lty = "dashed", alpha = 0.5) + -->
<!--   facet_wrap(~name, scales = "free") + -->
<!--   labs(x = "Parameter value", y = "Density") -->
<!-- ``` -->
``` {r}
res$final$plot()
```

### Posterior parameter distributions
Posterior parameter density estimates based on the mean and median icen.

```{r}
pmean = res$final$postMean %>%
  mutate(type = "Mean")

pmedian = res$final$postMed %>%
  mutate(type = "Median")

pmean %>% 
  bind_rows(pmedian) %>%
  pivot_longer(cols = -c(id, type)) %>%
  left_join(par_lims, by = "name") %>%
  ggplot(aes(x = value)) +
  geom_density(aes(col = type, fill = type), alpha = 0.5) +
  geom_vline(aes(xintercept = lower), 
             lty = "dashed", alpha = 0.5) +
  geom_vline(aes(xintercept = upper), 
             lty = "dashed", alpha = 0.5) +
  facet_wrap( ~ name, scales = "free") +
  labs(x = "Parameter value",
       y = "Density",
       fill = NULL,
       col = NULL) +
  theme(legend.position = "bottom")
```

### Table of support points

The following table lists the final support points, in order of decreasing probability
```{r}
res$final$popPoints %>% 
  arrange(desc(prob)) %>% 
  mutate(across(.cols = everything(), .fns = function(x) {
    round(x, digits = 4)
  })) %>% 
  mutate(`Cumulative Probability` = cumsum(prob)) %>% 
  rename(Probability = prob) %>% 
  datatable(options = list(pageLength = 10, searching = FALSE))
```

## Cycle information
```{r fig.dim = c(10, 6)}
res$cycle$plot()
```

<!-- THE FOLLOWING SECTION CONTAINS JAVASCRPT SUPPORT FUNCTIONS -->
```{js}
// This code ensures that click "Mean" tab changes all tabs to "Mean"
// Original code from https://community.rstudio.com/t/rmarkdown-html-change-all-tabs-on-click/160943


<!-- // make sure html is fully loaded -->
<!-- document.addEventListener("DOMContentLoaded", function() { -->
<!-- document.addEventListener('click', function(event) { -->
<!-- if (event.target.getAttribute('data-toggle') === 'tab') { -->
<!-- var clickedTabId = event.target.getAttribute('href').replace(/-[0-9]*/g, ''); -->

<!-- var tabs = document.querySelectorAll('[href^="' +                            clickedTabId + '"][data-toggle="tab"]') -->

<!-- for (var i = 0; i < tabs.length; i++) { -->
<!-- $(tabs[i]).tab('show'); -->
<!-- } -->
<!-- } -->
<!-- }) -->
<!-- }); -->
<!-- ``` -->



```{js}
$("input.hideshow").each(function(index, button) {
button.value = 'Hide Output';
$(button).click(function() {
var target = this.nextSibling ? this : this.parentNode;
target = target.nextSibling.nextSibling.nextSibling.nextSibling;
if (target.style.display == 'block' || target.style.display == '') {
target.style.display = 'none';
this.value = 'Show Output';
} else {
target.style.display = 'block';
this.value = 'Hide Output';
}
});
});
$("input.hideshow").click()
```

