---
title: "Pmetrics - report"
author: "Laboratory of Applied Pharmacokinetics and Bioinformatics (LAPKB)"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
params:
  res: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

library(tidyverse)
library(patchwork)
library(DT)
library(Pmetrics)
library(htmltools)
```


```{r setup-extra}
#res = params$res
setwd("~/LAPK/Pmetrics/Markus/Runs")
res <- PM_load(67)
```

## Summary

```{r}
did_converge = ifelse(res$NPdata$converge == 1, "converged", "did not converge")
```


The model includes `r res$NPdata$nsub` subjects, and a total of `r  sum(res$NPdata$nobs, na.rm = TRUE)` observations.

After `r res$NPdata$icyctot`/`r res$NPdata$icycmax` cycles, the model `r did_converge`.

## Performance metrics {.tabset}

```{r}
#summary function
metrics <- function(icen, outeq){
  pe <- res$op$summary(icen = icen, outeq = outeq)$pe %>%
    rename(`Mean PE` = mpe,
           `Mean weighted PE` = mwpe,
           `Mean square PE` = mspe,
           `Mean weighted squared PE` = mwspe,
           `Root mean squared PE` = rmse,
           `Root mean squared PE (%)` = percent_rmse,
           `Bias-adjusted mean squared PE` = bamspe,
           `Bias-adjusted mean weighted squared PE` = bamwspe
    ) %>% 
    mutate(across(.cols = everything(),
                  .fns = function(x) {
                    round(x, digits = 3)
                  })) %>% 
    t() %>% 
    datatable(
      colnames = c("Metric", "Value"),
      options = list(
        searching = FALSE,
        pageLength = 10,
        lengthChange = FALSE,
        info = FALSE,
        dom = "t",
        autoHideNavigation = TRUE
      )
    )  
  return(pe)
}


```

```{r, results = "asis"}

for(i in 1:res$NPdata$numeqt){
  cat("### Outeq ",i,"{.tabset .tabset-pills}\n")
  cat("\n")
  cat("#### Median\n")
  cat("\n")
  print(tagList(metrics("median", i)))
  cat("\n")
  cat("#### Mean\n")
  cat("\n")
  print(tagList(metrics("mean", i)))
  cat("\n")
}

```


## Observed-predicted plot {.tabset .tabset-pills}

Points should appear close to, and randomly distributed around, the (dashed) identity-line.

``` {r include = FALSE}

template <- c(
  "### Outeq {{outeq}} {.tabset .tabset-pills}\n",
  "#### Population {.tabset .tabset-pills}\n",
  "##### Median\n",
  "```{r message = F}\n",
  "res$op$plot(icen = 'median', pred.type = 'pop', outeq = {{outeq}},\n
  stats = list(x = 1, y = 0.1, font = list(size = 9)))\n",
  "```\n",
  "##### Mean\n",
  "```{r message = F}\n",
  "res$op$plot(icen = 'mean', pred.type = 'pop', outeq = {{outeq}},\n
  stats = list(x = 1, y = 0.1, font = list(size = 9)))\n",
  "```\n",
  "#### Posterior {.tabset .tabset-pills}\n",
  "##### Median\n",
  "```{r message = F}\n",
  "res$op$plot(icen = 'median', pred.type = 'post', outeq = {{outeq}},\n
  stats = list(x = 1, y = 0.1, font = list(size = 9)))\n",
  "```\n",
  "##### Mean\n",
  "```{r message = F}\n",
  "res$op$plot(icen = 'mean', pred.type = 'post', outeq = {{outeq}},\n
  stats = list(x = 1, y = 0.1, font = list(size = 9)))\n",
  "```\n",
  "\n"
)

numeqt <- max(res$op$data$outeq)
p <- lapply(1:numeqt, function(i) {
  knitr::knit_expand(
    outeq = i,
    text = template)
})

```
`r knitr::knit_child(text = unlist(p))`


## Posterior residuals by time {.tabset .tabset-pills}

Residuals should be randomly distributed around zero, with no apparent trend.
For how to interpret these plots, please review our [website](https://lapkb.github.io/Pmetrics). Weighted prediction errors are weighted by the observation standard deviation, i.e. the error polynomial.

```{r}

resid_plot <- function(x, y, icen, outeq){
  p <- res$op$data %>% 
    filter(icen == !!icen, outeq == !!outeq, pred.type == "post") %>% # Posterior concentrations (not pop)
    ggplot(aes(x = {{x}}, y = {{y}})) +
    geom_abline(slope = 0, intercept = 0, linetype = "dashed") +
    geom_smooth(method = "loess", se = FALSE, col = "black") +
    geom_point(pch = 21, col = "black", fill = "red") +
    theme_classic() +
    labs(x = "Time", y = ifelse(deparse(substitute(y))=="d", "Prediction error", "Weighted Prediction Error"))
  return(p)
}

```

```{r, results = "asis"}

for(i in 1:res$NPdata$numeqt){
  p1 <- resid_plot(time, d, "median", i)
  p2 <- resid_plot(time, wd, "median", i)
  p3 <- resid_plot(time, d, "mean", i)
  p4 <- resid_plot(time, wd, "mean", i)
  cat("\n")
  cat("### Outeq ",i,"{.tabset}\n")
  cat("\n")
  cat("#### Median\n")
  cat("\n\n")
  print(p1 / p2)
  cat("\n\n")
  cat("#### Mean\n")
  cat("\n\n")
  print(p3 / p4)
  cat("\n\n")
}


```


## Posterior residuals by concentration {.tabset}

Residuals should be randomly distributed around zero, with no apparent trend.
For how to interpret these plots, please review our [website](https://lapkb.github.io/Pmetrics). Weighted prediction errors are weighted by the observation standard deviation, i.e. the error polynomial.

```{r, results = "asis"}

for(i in 1:res$NPdata$numeqt){
  p1 <- resid_plot(obs, d, "median", i)
  p2 <- resid_plot(obs, wd, "median", i)
  p3 <- resid_plot(obs, d, "mean", i)
  p4 <- resid_plot(obs, wd, "mean", i)
  cat("\n")
  cat("### Outeq ",i,"{.tabset}\n")
  cat("\n")
  cat("#### Median\n")
  cat("\n\n")
  print(p1 / p2)
  cat("\n\n")
  cat("#### Mean\n")
  cat("\n\n")
  print(p3 / p4)
  cat("\n\n")
}


```


## Parameter values and distributions

The model included the following primary parameters
```{r}
par_lims = data.frame(
  name = head(names(res$final$popPoints), -1),
  lower = res$final$ab[,1],
  upper = res$final$ab[,2]
)

par_lims %>% 
  rename(`Parameter` = name,
         `Lower limit` = lower,
         `Upper limit` = upper) %>% 
  datatable(
    options = list(
      searching = FALSE,
      pageLength = 10,
      lengthChange = FALSE,
      info = FALSE,
      dom = "t",
      autoHideNavigation = TRUE
    )
  )
```

#### Marginal plots
Support points, height correspond to probability

```{r}
res$final$popPoints %>% 
  pivot_longer(cols = -prob) %>% 
  left_join(par_lims, by = "name") %>% 
  ggplot(aes(x = value, y = prob)) +
  geom_segment(aes(xend = value, yend = 0)) +
  geom_vline(aes(xintercept = lower), lty = "dashed", alpha = 0.5) +
  geom_vline(aes(xintercept = upper), lty = "dashed", alpha = 0.5) +
  facet_wrap(~name, scales = "free") +
  labs(x = "Parameter value", y = "Density")
```


#### Posterior parameter distributions
Posterior parameter density estimates based on the mean and median icen.

```{r}
pmean = res$final$postMean %>%
  mutate(type = "Mean")

pmedian = res$final$postMed %>%
  mutate(type = "Median")

pmean %>% 
  bind_rows(pmedian) %>%
  pivot_longer(cols = -c(id, type)) %>%
  left_join(par_lims, by = "name") %>%
  ggplot(aes(x = value)) +
  geom_density(aes(col = type, fill = type), alpha = 0.5) +
  geom_vline(aes(xintercept = lower), 
             lty = "dashed", alpha = 0.5) +
  geom_vline(aes(xintercept = upper), 
             lty = "dashed", alpha = 0.5) +
  facet_wrap( ~ name, scales = "free") +
  labs(x = "Parameter value",
       y = "Density",
       fill = NULL,
       col = NULL) +
  theme(legend.position = "bottom")
```

#### Table of support points

The following table lists the final support points, in order of decreasing probability
```{r}
res$final$popPoints %>% 
  arrange(desc(prob)) %>% 
  mutate(across(.cols = everything(), .fns = function(x) {
    round(x, digits = 4)
  })) %>% 
  mutate(`Probability (cumulative sum)` = cumsum(prob)) %>% 
  rename(Probability = prob) %>% 
  datatable(options = list(pageLength = 10, searching = FALSE))
```



## Cycle information
```{r}
cycle_info = data.frame(
  cycle = res$cycle$data$cycnum,
  ll = res$cycle$data$ll,
  AIC = res$cycle$data$aic,
  BIC = res$cycle$data$bic,
  gamlam = res$cycle$data$gamlam
  
)

cycle_info = slice_tail(cycle_info, prop = 0.8)

plot_ll = cycle_info %>% 
  ggplot(aes(x = cycle, y = ll)) +
  geom_line() +
  labs(subtitle = "-2 x log likelihood",
       x = "Cycle number", y = NULL) +
  theme_classic() +
  ggpubr::grids() +
  theme(legend.position = "bottom")

plot_aic_bic = cycle_info %>% 
  pivot_longer(cols = c(AIC, BIC)) %>% 
  ggplot(aes(x = cycle, y = value, col = name)) +
  geom_line() +
  labs(subtitle = "AIC/BIC",
       x = "Cycle number", y = NULL, col = NULL) +
  theme_classic() +
  ggpubr::grids() +
  theme(legend.position = "bottom")

plot_gamlam = cycle_info %>% 
  ggplot(aes(x = cycle, y = gamlam)) +
  geom_line() +
  labs(subtitle = "Gamma / Lambda",
       x = "Cycle number", y = NULL) +
  theme_classic() +
  ggpubr::grids() +
  theme(legend.position = "bottom")

plot_mean = res$cycle$data$mean %>% 
  data.frame() %>% 
  mutate(cycle = res$cycle$data$cycnum) %>% 
  slice_tail(prop = 0.8) %>% 
  pivot_longer(cols = -cycle) %>% 
  ggplot(aes(x = cycle, y = value, col = name)) +
  geom_line() + 
  labs(subtitle = "Normalized mean",
       x = "Cycle number", y = NULL, col = NULL) +
  theme_classic() +
  ggpubr::grids() +
  theme(legend.position = "bottom")

plot_sd = res$cycle$data$sd %>% 
  data.frame() %>% 
  mutate(cycle = res$cycle$data$cycnum) %>% 
  slice_tail(prop = 0.8) %>% 
  pivot_longer(cols = -cycle) %>% 
  ggplot(aes(x = cycle, y = value, col = name)) +
  geom_line() + 
  labs(subtitle = "Normalized standard deviation",
       x = "Cycle number", y = NULL, col = NULL) +
  theme_classic() +
  ggpubr::grids() +
  theme(legend.position = "bottom")

plot_median = res$cycle$data$median %>% 
  data.frame() %>% 
  mutate(cycle = res$cycle$data$cycnum) %>% 
  slice_tail(prop = 0.8) %>% 
  pivot_longer(cols = -cycle) %>% 
  ggplot(aes(x = cycle, y = value, col = name)) +
  geom_line() + 
  labs(subtitle = "Normalized median",
       x = "Cycle number", y = NULL, col = NULL) +
  theme_classic() +
  ggpubr::grids() +
  theme(legend.position = "bottom")

plot_design = "
AB
CD
EF
"

plot_ll + plot_aic_bic + plot_gamlam + plot_mean + plot_sd + plot_median +
  plot_layout(design = plot_design, guides = "collect") &
  theme(legend.position = "bottom")
```

<!-- THE FOLLOWING SECTION CONTAINS JAVASCRPT SUPPORT FUNCTIONS -->
```{js}
// This code ensures that click "Mean" tab changes all tabs to "Mean"
// Original code from https://community.rstudio.com/t/rmarkdown-html-change-all-tabs-on-click/160943


<!-- // make sure html is fully loaded -->
<!-- document.addEventListener("DOMContentLoaded", function() { -->
<!-- document.addEventListener('click', function(event) { -->
<!-- if (event.target.getAttribute('data-toggle') === 'tab') { -->
<!-- var clickedTabId = event.target.getAttribute('href').replace(/-[0-9]*/g, ''); -->

<!-- var tabs = document.querySelectorAll('[href^="' +                            clickedTabId + '"][data-toggle="tab"]') -->

<!-- for (var i = 0; i < tabs.length; i++) { -->
<!-- $(tabs[i]).tab('show'); -->
<!-- } -->
<!-- } -->
<!-- }) -->
<!-- }); -->
<!-- ``` -->



```{js}
$("input.hideshow").each(function(index, button) {
button.value = 'Hide Output';
$(button).click(function() {
var target = this.nextSibling ? this : this.parentNode;
target = target.nextSibling.nextSibling.nextSibling.nextSibling;
if (target.style.display == 'block' || target.style.display == '') {
target.style.display = 'none';
this.value = 'Show Output';
} else {
target.style.display = 'block';
this.value = 'Hide Output';
}
});
});
$("input.hideshow").click()
```

