---
title: "Pmetrics - Report"
author: "Laboratory of Applied Pharmacokinetics and Bioinformatics (LAPKB)"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
params:
  res: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(dplyr)
```


```{r setup-extra}
res = params$res
#res = Pmetrics::ITex
```

## Summary

```{r}
all_res <- nsub <- nobs <- numeqt <- icyctot <- icycmax <- NULL #declare for CRAN check

if(is.null(res$ITdata)){
  all_res <- res$NPdata
  np <- TRUE
} else {
  all_res <- res$ITdata
  np <- FALSE
}
did_converge <- ifelse(all_res$converge == 1, "converged.", "did not converge.")
nsub <- all_res$nsub
nobs <- sum(all_res$nobs, na.rm = TRUE)
numeqt <- all_res$numeqt
icyctot <- all_res$icyctot
icycmax <- all_res$icycmax
```


The data had `r nsub` subjects and a total of `r  nobs` observations.

After `r paste0(icyctot, " of ", icycmax, " cycles, the model ", did_converge)`

## Observed-predicted plot {.tabset .tabset-pills}

Points should appear close to, and randomly distributed around, the (dashed) identity-line.

``` {r include = FALSE}

template <- c(
  "### Outeq {{outeq}} {.tabset .tabset-pills}\n",
  "#### Population {.tabset .tabset-pills}\n",
  "##### Median\n",
  "```{r message = F}\n",
  "res$op$plot(icen = 'median', pred.type = 'pop', outeq = {{outeq}},\n
  stats = list(x = 1, y = 0.1, font = list(size = 9)))\n",
  "```\n",
  "##### Mean\n",
  "```{r message = F}\n",
  "res$op$plot(icen = 'mean', pred.type = 'pop', outeq = {{outeq}},\n
  stats = list(x = 1, y = 0.1, font = list(size = 9)))\n",
  "```\n",
  "#### Posterior {.tabset .tabset-pills}\n",
  "##### Median\n",
  "```{r message = F}\n",
  "res$op$plot(icen = 'median', pred.type = 'post', outeq = {{outeq}},\n
  stats = list(x = 1, y = 0.1, font = list(size = 9)))\n",
  "```\n",
  "##### Mean\n",
  "```{r message = F}\n",
  "res$op$plot(icen = 'mean', pred.type = 'post', outeq = {{outeq}},\n
  stats = list(x = 1, y = 0.1, font = list(size = 9)))\n",
  "```\n",
  "\n"
)

numeqt <- max(res$op$data$outeq)
p <- lapply(1:numeqt, function(i) {
  knitr::knit_expand(
    outeq = i,
    text = template)
})

```
`r knitr::knit_child(text = unlist(p))`

## Prediction metrics {.tabset .tabset-pills}

*Measures of bias. Others are measures of imprecision.

^Same units as observations and predictions. Others are squared or dimensionless.
```{r}
#summary function
metrics <- function(icen, pred.type, outeq){
  pe <- res$op$summary(icen = icen,  pred.type = pred.type, outeq = outeq)$pe %>%
    dplyr::rename(`Mean PE (prediction - observation)*^` = mpe,
                  `Mean weighted PE*^` = mwpe,
                  `Mean squared PE` = mspe,
                  `Mean weighted squared PE` = mwspe,
                  `Root mean squared PE^` = rmse,
                  `Root mean squared PE (%)` = percent_rmse,
                  `Bias-adjusted mean squared PE` = bamspe,
                  `Bias-adjusted mean weighted squared PE` = bamwspe
    ) %>% 
    dplyr::mutate(across(everything(), \(x) round(x, 3))) %>% 
    t() %>% 
    DT::datatable(
      colnames = c("Metric", "Value"),
      options = list(
        searching = FALSE,
        pageLength = 10,
        lengthChange = FALSE,
        info = FALSE,
        dom = "t",
        autoHideNavigation = TRUE
      )
    )  
  return(pe)
}


```

```{r include = FALSE}

template <- c(
   "### Outeq {{outeq}} {.tabset .tabset-pills}\n",
  "#### Population {.tabset .tabset-pills}\n",
  "##### Median\n",
  "```{r message = F}\n",
  "metrics(icen = 'median', pred.type = 'pop', outeq = {{outeq}})\n",
  "```\n",
  "##### Mean\n",
  "```{r message = F}\n",
  "metrics(icen = 'mean', pred.type = 'pop', outeq = {{outeq}})\n",
  "```\n",
  "#### Posterior {.tabset .tabset-pills}\n",
  "##### Median\n",
  "```{r message = F}\n",
  "metrics(icen = 'median', pred.type = 'post', outeq = {{outeq}})\n",
  "```\n",
  "##### Mean\n",
  "```{r message = F}\n",
  "metrics(icen = 'mean', pred.type = 'post', outeq = {{outeq}})\n",
  "```\n",
  "\n"
)


numeqt <- max(res$op$data$outeq)
p <- lapply(1:numeqt, function(i) {
  knitr::knit_expand(
    outeq = i,
    text = template)
})


```
`r knitr::knit_child(text = unlist(p))`

## Prediction residuals {.tabset .tabset-pills}

Residuals should be randomly distributed around zero, with no apparent trend.
For how to interpret these plots, please review our [website](https://lapkb.github.io/Pmetrics). Weighted prediction errors are weighted by the observation standard deviation, i.e. the error polynomial.

``` {r include = FALSE}

template <- c(
  "### Outeq {{outeq}} {.tabset .tabset-pills}\n",
  "#### Population {.tabset .tabset-pills}\n",
  "##### Median\n",
  "```{r message = F}\n",
  "res$op$plot(resid = T, icen = 'median', pred.type = 'pop', outeq = {{outeq}})\n",
  "```\n",
  "##### Mean\n",
  "```{r message = F}\n",
  "res$op$plot(resid = T, icen = 'mean', pred.type = 'pop', outeq = {{outeq}})\n",
  "```\n",
  "#### Posterior {.tabset .tabset-pills}\n",
  "##### Median\n",
  "```{r message = F}\n",
  "res$op$plot(resid = T, icen = 'median', pred.type = 'post', outeq = {{outeq}})\n",
  "```\n",
  "##### Mean\n",
  "```{r message = F}\n",
  "res$op$plot(resid = T, icen = 'mean', pred.type = 'post', outeq = {{outeq}})\n",
  "```\n",
  "\n"
)

numeqt <- max(res$op$data$outeq)
p <- lapply(1:numeqt, function(i) {
  knitr::knit_expand(
    outeq = i,
    text = template)
})

```
`r knitr::knit_child(text = unlist(p), options = list(fig.width = 10))`

## Parameter values and distributions

### Parameter summmaries
The model included the following primary parameters
```{r}
par_lims <- data.frame(
  name = names(res$final$popMean),
  lower = res$final$ab[,1],
  upper = res$final$ab[,2],
  ave = t(res$final$popMean),
  med = t(res$final$popMedian),
  sd = t(res$final$popSD),
  cv = t(res$final$popCV),
  shrink = 100*t(res$final$shrinkage)
) %>%
  dplyr::mutate(across(4:8,\(x) round(x, 3)))

par_lims %>% 
  dplyr::rename(`Parameter` = name,
                `Lower limit` = lower,
                `Upper limit` = upper,
                `Mean` = ave,
                `Median` = med,
                `SD` = sd,
                `CV%` = cv,
                `%shrink` = shrink) %>% 
  DT::datatable(
    options = list(
      searching = FALSE,
      pageLength = 10,
      lengthChange = FALSE,
      info = FALSE,
      dom = "t",
      autoHideNavigation = TRUE
    ),
    rownames = FALSE
  )
```

### Marginal plots
Probability distributions, dashed lines are limits.

``` {r}
res$final$plot()
```

<!-- ### Posterior parameter distributions -->
<!-- Posterior parameter density estimates based on the mean and median icen. -->

<!-- ```{r} -->
<!-- pmean = res$final$postMean %>% -->
<!--   mutate(type = "Mean") -->

<!-- pmedian = res$final$postMed %>% -->
<!--   mutate(type = "Median") -->

<!-- pmean %>%  -->
<!--   bind_rows(pmedian) %>% -->
<!--   pivot_longer(cols = -c(id, type)) %>% -->
<!--   left_join(par_lims, by = "name") %>% -->
<!--   ggplot(aes(x = value)) + -->
<!--   geom_density(aes(col = type, fill = type), alpha = 0.5) + -->
<!--   geom_vline(aes(xintercept = lower),  -->
<!--              lty = "dashed", alpha = 0.5) + -->
<!--   geom_vline(aes(xintercept = upper),  -->
<!--              lty = "dashed", alpha = 0.5) + -->
<!--   facet_wrap( ~ name, scales = "free") + -->
<!--   labs(x = "Parameter value", -->
<!--        y = "Density", -->
<!--        fill = NULL, -->
<!--        col = NULL) + -->
<!--   theme(legend.position = "bottom") -->
<!-- ``` -->

```{r, conditional-text, eval = np, echo = FALSE}
knitr::asis_output("### Table of support points\n\n
The following table lists the final support points, in order of decreasing probability.\n")
```

```{r}
if(np){
  res$final$popPoints %>% 
    dplyr::arrange(desc(prob)) %>% 
    dplyr::mutate(across(everything(), \(x) round(x, 3)
    )) %>% 
    dplyr::mutate(`Cumulative Probability` = cumsum(prob)) %>% 
    dplyr::rename(Probability = prob) %>% 
    DT::datatable(options = list(pageLength = 10, searching = FALSE))
}
```

## Cycle information
```{r fig.dim = c(10, 6)}
res$cycle$plot()
```

<!-- THE FOLLOWING SECTION CONTAINS JAVASCRPT SUPPORT FUNCTIONS -->
```{js}
// This code ensures that click "Mean" tab changes all tabs to "Mean"
// Original code from https://community.rstudio.com/t/rmarkdown-html-change-all-tabs-on-click/160943


<!-- // make sure html is fully loaded -->
<!-- document.addEventListener("DOMContentLoaded", function() { -->
<!-- document.addEventListener('click', function(event) { -->
<!-- if (event.target.getAttribute('data-toggle') === 'tab') { -->
<!-- var clickedTabId = event.target.getAttribute('href').replace(/-[0-9]*/g, ''); -->

<!-- var tabs = document.querySelectorAll('[href^="' +                            clickedTabId + '"][data-toggle="tab"]') -->

<!-- for (var i = 0; i < tabs.length; i++) { -->
<!-- $(tabs[i]).tab('show'); -->
<!-- } -->
<!-- } -->
<!-- }) -->
<!-- }); -->
<!-- ``` -->



```{js}
$("input.hideshow").each(function(index, button) {
button.value = 'Hide Output';
$(button).click(function() {
var target = this.nextSibling ? this : this.parentNode;
target = target.nextSibling.nextSibling.nextSibling.nextSibling;
if (target.style.display == 'block' || target.style.display == '') {
target.style.display = 'none';
this.value = 'Show Output';
} else {
target.style.display = 'block';
this.value = 'Hide Output';
}
});
});
$("input.hideshow").click()
```

