---
title: "Data objects"
output: 
  rmarkdown::html_vignette:
    css: style.css
vignette: >
  %\VignetteIndexEntry{Data objects}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo=F, message=F}
library(Pmetrics)
```

## Introduction

Pmetrics data objects are either R6 objects loaded into memory or files, usually comma-separated (.csv). It is possible
to use other separators, like the semicolon, by setting the appropriate argument with `setPMoptions()`. 

```{r echo = T, eval = F}
setPMoptions(sep = ";", dec = ",")
# changes field separator to ";" from default ","
# and decimal separator from "." to ","
```

Examples of programs that can save .csv files are any text editor (e.g. TextEdit on Mac, Notepad on Windows) or spreadsheet program (e.g. Excel). 

## Data file use

<span class="r6">R6</span> 

R6 Pmetrics introduces a new concept, the data object. The idea of this object is to represent a dataset that is going to be modeled/simulated.
All its behaviour is represented by the class `PM_data`. This class allows datasets to be checked, plotted, written to disk and more. Use `PM_data$new("filename")` to create a `PM_data()` object by reading the file.

```{r echo=T, eval=FALSE}
# assume that data.csv is in the working directory
data1 <- PM_data$new("data.csv")
```

You can also build an appropriate data frame in R and provide that as an argument to `PM_data$new()`.

```{r echo=T, eval=FALSE}
# assume df is data frame with at least these columns:
# id, time, dose, out
data1 <- PM_data$new(df)
```


Lastly, you can take advantage of the `addEvent` method in `PM_data` objects to build a data object on the fly. This can be particularly useful for making quick simulation templates. Start with an empty call to `PM_data$new()` and add successive rows. See [PM_data] for details under the `addEvent` method.

```{r echo=TRUE, eval=TRUE}
dat <- PM_data$new()$addEvent(id = 1, time = 0, dose = 100, addl = 5, ii = 24)$addEvent(id = 1, time = 144, out = -1)
dat$addEvent(id = 1, wt = 75, validate = TRUE)
```

*Notes on the above statments:*

1. Lack of time element in the last `addEvent` will add *wt=75* to all rows for id = 1
2. Use `validate = TRUE` as an argument in the last `addEvent` to finalize creation
3. Note that you can chain `addEvent` as shown in the first statement and also update existing `PM_data` objects as shown in the second statement.


Here's what `dat$data` and `dat$standard_data` look like in the viewer.
``` {r echo = FALSE, eval = TRUE}
knitr::kable(dat$data)
knitr::kable(dat$standard_data)
```

Once you have created the `PM_data` object, you never need to create it again during your R session. You also don't have to bother copying the data file to the Runs folder each time you run the model, like you used to do with older **("Legacy")** versions of Pmetrics. The data are stored in memory and can be used in any Pmetrics function that needs it.


## Data format

<span class="r6">R6</span>

R6 Pmetrics can use file or data frame input. The  format is very flexible. The only required columns are those below. Unlike Legacy Pmetrics, there are no requirements for a header or to prefix the ID column with "\#". However, any subsequent row
that begins with "\#" will be ignored, which is helpful if you want to
exclude data from the analysis, but preserve the integrity of the
original dataset, or to add comment lines. The column order can be anything you wish, but the names should be the same as below. Ultimately, `PM_data$new()` converts all valid data into a standardized format, which
is the same as used in Legacy Pmetrics.

* ***ID*** This field can be numeric or character and
identifies each individual. All rows must contain an ID, and all
records from one individual must be contiguous. IDs may be any alphanumeric combination. 
Unlike Legacy Pmetrics, the number of subjects is unlimited in Rust.

* ***TIME*** This is the elapsed time in decimal hours since the first
event. You may also specify time as clock time if you include a ***DATE*** column below. In that case the default format is HH:MM. Other formats can be specified. See `PM_data()` for more details. Every row must have an entry, and within a given ID, rows must be sorted chronologically, earliest to latest.
 
* ***DATE***  This column is only required if ***TIME*** is clock time, detected by the presence of ":". The default format of the date column is YYYY-MM-DD. As for ***TIME***, other formats can be specified. See `?PM_data` for more details.

* ***DOSE*** This is the dose amount. It should be "." for observation rows.

* ***OUT*** This is the observation, or output value, e.g., concentration. For dose rows, it should be ".". If an observation occurred at a given time, but the result was missing, e.g. sample lost or below the limit of quantification, this should be coded as -99.

* ***COVARIATES\...*** Columns with names other than the above will be interpreted as covariates.

When `PM_data()` reads a file, it will standardize it to the format below. This means some inferences are made. For example, in the absence of `EVID`, all doses are interpreted as oral. If they are infusions, `DUR` must be included. `EVID` only needs to be included if `EVID=4` (reset event) is required, described below. Similarly, `INPUT` and `OUTEQ` are only required if multiple inputs or outputs are being modeled. Lastly, `ADDL` and `II` are optional.

Lastly, the standardized data are checked for errors and if found, Pmetrics 
generates a report with the errors and will attempt to fix those that it can. 


### Standardized Data ###

The order, capitalization and names of the header and the first 12 columns are fixed. All entries must be numeric, with the
exception of `ID` and "." for non-required placeholder entries. Any subsequent row
that begins with "\#" will be ignored, as above.

A full example data file is below, with details following.


```{r echo=F, results='asis'}
tab <- read.csv("Data/mdata.csv")
tab$OUT <- as.character(tab$OUT)
knitr::kable(tab)
```

* ***ID*** This field can be numeric or character and
identifies each individual. All rows must contain an ID, and all
records from one individual must be contiguous. Any subsequent row
that begins with "\#" will be ignored, which is helpful if you want to
exclude data from the analysis, but preserve the integrity of the
original dataset, or to add comment lines

* ***EVID*** This is the event ID field. It can be 0, 1, or 4. Every row
must have an entry.

    + 0 = observation

    + 1 = input (e.g. dose)

    + 2, 3 are currently unused

    + 4 = reset, where all compartment values are set to 0 and the time
counter is reset to 0. This is useful when an individual has multiple
sampling episodes that are widely spaced in time with no new
information gathered. This is a dose event, so dose information needs
to be complete. The `TIME` value for `EVID=4` should be 0, and subsequent
rows should increase monotonically from 0 until the last record or until
another `EVID=4` event, which will restart time at 0.

* ***DATE*** This is optional, and only required if `TIME` (below) is
clock time rather than decimal time. Pmetrics will attempt to detect the
date format by examining all the records to find the most unambiguous format.
However, the format can be specified in the `PM_data()` function. If present,
every row must have an entry, and rows within a given `ID` should be sorted
by date, earliest to latest.

* ***TIME*** This is the elapsed time in decimal hours since the first
event. If `TIME` is clock time, there must be a `DATE` column. As for `DATE`,
Pmetrics will attempt to understand the clock time format, and the format can
be unambiguously specified in the `PM_data()` function.
Every row must have an entry, and within a given `ID`, rows
must be sorted chronologically, earliest to latest.

* ***DUR*** This is the duration of an infusion in hours. If `EVID=1`,
there must be an entry, otherwise it is ignored and should have a "." placeholder. 
For a bolus (e.g. an oral dose), set the value equal to 0.

* ***DOSE*** This is the dose amount. If `EVID=1`, there must be an entry,
otherwise it is ignored and should have a "." placeholder. The first event
at `TIME=0` for a given `ID` must be a dose event, even if the dose amount is 0.

* ***ADDL*** This specifies the number of additional doses to give at
interval `II`. `ADDL` can be positive or negative.
If positive, it is the number of doses to give after the dose at time 0.
If negative, it is the number of doses to give before the dose at time 0.
It may be missing (".") for dose events (`EVID=1` or `EVID=4`), in which
case it is assumed to be 0. It is ignored for observation (`EVID=0`)
events. Be sure to adjust the time entry for the subsequent row, if
necessary, to account for the extra doses. All compartments in the
model will contain the predicted amounts of drug at the end of the
`II` interval after the last `ADDL` dose.

* ***II*** This is the interdose interval and is only relevant if `ADDL`
is not equal to 0, in which case `II` cannot be missing. If `ADDL=0` or is
missing, II is ignored.

* ***INPUT*** This defines which input (i.e. drug) the `DOSE` corresponds
to. The model defines which compartments receive the input(s). If only modeling
one drug, INPUT is unnecessary, as all values will be assumed to be 1.

* ***OUT*** This is the observation, or output value, and it is
always required. If `EVID=0`, there
must be an entry. For such events, if the observation is missing,
e.g. a sample was lost or not obtained, this must be coded as -99. 
It will be
ignored for any other `EVID` and therefore can be ".". 

* ***OUTEQ*** This is the output equation number that corresponds to the
OUT value. Output equations are defined in the model file. If only modeling 
one output, this column is unnecessary, as all values will be assumed to be 1.

* ***C0, C1, C2, C3*** These are the coefficients for the assay error
polynomial for that observation. Each subject may have up to one set
of coefficients per output equation. If more than one set is detected
for a given subject and output equation, the last set will be used. If
there are no available coefficients, these cells may be omitted. If they
are included, for events which are not observations, they can be filled
with "." as a placeholder. In data fitting, if the coefficients are present
in the data file, Pmetrics will use them. If missing, Pmetrics will look for
coefficients defined in the model.

* ***COVARIATES***\... Any column named other than above is assumed
to be a covariate, one column per covariate. The first row for any subject
must have a value for all covariates, since the first row is always a dose.
**Covariates are handled differently than in Legacy Pmetrics.** In Legacy, they
were only considered at the times of dose events (`EVID=1`). In Pmetrics ≥ 3.0, they are
considered at all times, including observation events (`EVID=0`). Therefore,
to enter a new covariate value at a time other than a dose or an observation,
create a row at the appropriate time (and possibly date if using clock/calendar),
making the row either a dose row (`EVID=1`) with `DOSE=0` or an observation
row (`EVID=0`) with `OUT=-99` (missing). By default, covariate values are linearly
interpolated between entries. This is useful for covariates like weight, which 
may vary from measurement to measurement. You can change this behavior to make them piece-wise
constant, i.e. carried forward from the previous value until a new value causes an 
instant change. This could be used, for example, to indicate periods of off and on 
dialysis. See `vignette("models")` for more details.

## Manipulation of CSV files

<span class="r6">R6</span>

#### Read

As we have seen,  `PM_data$new("filename")` will 
read an appropriate data file in the current working directory to create a 
new `PM_data()` object. 

#### Write

`PM_data$save("filename")` will write the `PM_data()` object to a file called
"filename". This can be useful if you have loaded or created a data file and then
changed it in R.

#### Standardize

`PM_data$new()` automatically standardizes the data into the full format. This includes conversion of calendar date / clock time into decimal elapsed time.

#### Validate

`PM_data$new()` automatically calls `PMcheck()` so the data are validated as the data object is created.
  
#### Data conversion

* `PMwrk2csv()` This function will convert old-style, single-drug USC\*PACK
.wrk formatted files into Pmetrics data .csv files. 

* `PMmb2csv()` This function will convert USC\*PACK .mb files into Pmetrics
data .csv files.

* `NM2PM()` Although the structure of Pmetrics data files is similar to
NONMEM, there are some differences. This function attempts to
automatically convert to Pmetrics format. It has been tested on several
examples, but there are probably NONMEM files which will cause it to
crash. 

