---
title: "Introduction"
output: 
  rmarkdown::html_vignette:
    css: style.css
vignette: >
  %\VignetteIndexEntry{Introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo=F, message=F}
```

Thank you for your interest in Pmetrics! This guide provides
instructions and examples to assist users of the Pmetrics R package, by
the Laboratory of Applied Pharmacokinetics and Bioinformatics at the
Saban Research Institute, Children's Hospital Los Angeles, and the
Department of Pediatrics, Keck School of Medicine, University of
Southern California. Please see our website at
[[http://www.lapk.org]{.underline}](http://www.lapk.org) for more
information.


## R6 architecture

**As of v. 2.0**, Pmetrics shifted to an architecture less dependent on
reading and writing files. Data files are unchanged.

Model files can now be defined as an R object, instead of in a text file. 
The easiest way to accomplish this is with our model builder app, which is
described in the article on Models.

Pmetrics can suppoort legacy runs with the old-style text file models, but
users are encouraged to change to the new methods. Throughout this manual we
will indicate the new style with the <span class="r6">R6</span> designation to reflect the
object-oriented R6 style of programming available in R that makes it more
consistent with object-oriented languages such as Python. We will indicate
old-style approaches with <span class="legacy">Legacy</span>.

Functions and classes in <span class="r6">R6</span> are generally named with "PM_" as the prefix, e.g. `PM_load()` or `PM_final()`. In <span class="legacy">Legacy</span>, functions and classes are prefixed without the underscore "_", such as `PMload()`, and `PMfinal()`.

**Tips for using this guide.**

* Items that are hyperlinked can be selected to cross reference
within this manual or link to external sites.

* `Items` correspond to inline examples of R code, which are not evaluated
in this document, but serve as templates for what may be typed into your 
R console or script. They may not necessarily be executable if typed verbatim.

## Disclaimer

You, the user, assume all responsibility for acting on the results
obtained from Pmetrics. The Laboratory of Applied Pharmacokinetics and
Bioinformatics (LAPKB), members and consultants to LAPKB, and Children's
Hospital Los Angeles and the University of Southern California and their
employees assume no liability whatsoever. Your use of the package
constitutes your agreement to this provision.

## Installation

Pmetrics and all required components will run under Mac (Unix), Windows,
and Linux. Instructions on obtaining and installing the package are in `vignette("pmetrics")`.

## What This Manual Is Not

We assume that the user has familiarity with population modeling and R,
and thus this manual is not a tutorial for basic concepts and techniques
in either domain. We have tried to make the R code simple, regular and
well documented. A very good free online resource for learning the
basics of R can be found at [Stat Methods](http://www.statmethods.net/index.html).

We recognize that initial use of a new software package can be complex,
so please feel free to contact us at any time, preferably through the
[Pmetrics Disussions](https://github.com/LAPKB/Pmetrics/discussions). You can report [issues](https://github.com/LAPKB/Pmetrics/issues) online as well.

This manual is also not intended to be a theoretical treatise on the
algorithms used in IT2B or NPAG. For that, the user is directed to our
[website](http://www.lapk.org/publications.php).

## Getting Help and Updates


Within R, you can  use
`help("command")` or more simply just `?command` in the R console to see detailed help files for any Pmetrics command. Many commands have examples included in this documentation and you can execute the examples with `example(command)`.

You can also type `PMmanual()` to launch this website from within Pmetrics. 

Pmetrics will check for updates automatically every time you load it
with `library(Pmetrics)` and you are connected to the internet. 
If an update is available, it will provide a
brief message to inform you. You can then reinstall the package from github.

<!--When bugs arise in Pmetrics, you may see a start up message to inform you of 
the bug and a patch can be installed by the command `PMpatch()` if available. 
Note that patches must be reinstalled with this command every time you launch
Pmetrics, until the bug is corrected in the next version.
-->

<!--As of version 1.0.0 Pmetrics has graphical user interface (GUI)
capability for many functions. Using `PMcode("function")` will launch the
GUI in your default browser. While you are interacting with the GUI, R
is "listening" and no other activity is possible. The GUI is designed to
generate Pmetrics R code in response to your input in a friendly,
intuitive environment. That code can be copied and pasted into your
Pmetrics R script. You can also see live plot previews with the GUI. All
this is made possible with the 'shiny' package for R.


Currently, the following GUIs are available: `PMcode("run")` and
`PMcode("plot")`. 
--> 

## Customizing Options

You can change global options in Pmetrics with 
`setPMoptions()`. The `getPMoptions()` function will return the 
current options.