---
title: "Created Objects"
output: 
  rmarkdown::html_vignette:
    css: style.css
bibliography: Citations.bib
vignette: >
  %\VignetteIndexEntry{Created Objects}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo=F, message=F}
library(Pmetrics)
```

## Objects created after a run

After a successful NPAG run, an R datafile is saved in the
output subdirectory of the newly created numerically ordered folder in
the working directory. This file is called "PMout.Rdata". These data files can be loaded by ensuring that the
Runs folder is set as the working directory, and then using the Pmetrics
command `PM_load()`.

There are several Pmetrics data objects contained within the Rdata files
which are loaded, making these objects available for
plotting and other analysis. 


`PM_load()` returns a `PM_result` object. The fields in the object are in the table below.


```{r echo = F, results = 'asis'}
tabObj <- read.csv("Data/ObjectsR6.csv", na.strings = ".")
knitr::kable(tabObj)
```

To access the fields in a `PM_result` object, use the "$". 

```{r echo=T, eval = F}
res1 <- PM_load(1)
res1$op
res1$op$time
res1$op$data # this accesses the whole observed vs. predicted data frame
```

Most of the fields are themselves R6 objects with methods.

```{r echo = T, eval = F}
res1 <- PM_load(1)
res1$op
res1$op$summary()
res1$op$plot()
```


