---
title: "Overview"
output: 
  rmarkdown::html_vignette:
    css: style.css
vignette: >
  %\VignetteIndexEntry{Overview}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo=F, message=F}
library(Pmetrics)
```

## Software engines

There are three main software engines that Pmetrics controls.

<!-- * **IT2B** is the ITerative 2-stage Bayesian parametric population PK modeling -->
<!-- program. It is generally used to estimate parameter ranges to pass to -->
<!-- NPAG. It will estimate values for population model parameters under the -->
<!-- assumption that the underlying distributions of those values are normal -->
<!-- or transformed to normal, e.g. log normal. -->

* **NPAG** is the Non-parametric Adaptive Grid software. It will create a
non-parametric population model consisting of discrete support points,
each with a set of estimates for all parameters in the model plus an
associated probability (weight) of that set of estimates. There can be
at most one point for each subject in the study population. There is no
need for any assumption about the underlying distribution of model
parameter values.

* **NPOD** is the Non-parametric Optimal Design software. Like NPAG,
it will create a
non-parametric population model consisting of discrete support points,
each with a set of estimates for all parameters in the model plus an
associated probability (weight) of that set of estimates. While NPAG 
searches parameter hyperspace randomly, NPOD uses likelihood gradients
to search systematically. This usually results in faster convergence, 
but currently, NPOD is more likely to get stuck in a "local" minimum 
of the likelihood surface, i.e. it has a higher chance of not finding the
globally optimal support points. NPOD is under active development.

* The **Simulator** is a semi-parametric Monte Carlo simulation software
program that can use the output of NPAG of NPOD to build randomly
generated response profiles (e.g. time-concentration curves) for a given
population model, parameter estimates, and data input. Simulation from a
non-parametric joint density model, i.e. NPAG output, is possible, with
each point serving as the mean of a multivariate normal distribution,
weighted according to the weight of the point. The covariance matrix of
the entire set of support points is divided equally among the points for
the purposes of simulation.

## Pmetrics control functions

Pmetrics uses`PM_data()` to create data objects,  `PM_model()` to create model 
objects, and `PM_fit()` to create
objects that combine the model with the data, ready to be run (fitted),
generating probability distributions for primary model parameters. 
These are extensively
documented within R by using the `help(command)` or `?command` syntax.

These functions replace the following Legacy functions: `ITrun`, `ERRrun`, 
`NPrun`.

Invoking the simulator in R6 becomes a method attached to `PM_result()` objects
or by using `PM_sim$run()` for models, parameter value probability distributions
and template data not derived from a previous fit, e.g. when lifted from an article.


### Run functions

Once a `PM_fit()` object is created, which combines a model with a data file,
it can be run by using the syntax `$run()` to access the appropriate function
defined for the `PM_fit()` object. 

```{r echo=T, eval=FALSE}
fit1 <- PM_fit$new(data, model)
fit1$run(options)
```



For the Simulator, the `$sim` method for a `PM_result` object will execute the 
program directly within R and return a `PM_sim` object. 

```{r echo=T, eval=FALSE}
run1 <- PM_load(1)
sim1 <- run1$sim(data = "new.csv")
```

The first line loads previous results of run 1 into a `PM_result()` object called run1. The second line uses the model and prior in run1 with new data template.

An equivalent method is to run a new `PM_sim()` directly with `PM_sim$run()`. It
also returns a `PM_sim()` object, i.e. there is no longer any need to run `SIMparse`.

```{r echo=T, eval=FALSE}
sim1 <- PM_sim$run(poppar = list(...), model = "model.txt", data = "new.csv")
```

See `vignette("simulation")` for details on specifying `poppar` this way.


For the simulator the results are returned automatically to the object assigned
to contain the output of the simulation, e.g. `sim1` below. 

```{r echo=T, eval=FALSE}
sim1 <- PM_result$sim(...)
```

As mentioned above, there is no longer any
need to use the `SIMparse` Legacy function, because the 
`SIMrun` and `SIMparse` Legacy functions are combined within the `$sim()` method for `PM_result()` objects or the `$run()` method for `PM_sim()` objects.


### Saving functions

The Pmetrics R6 objects `PM_fit()`, `PM_result()`, `PM_sim()`, `PM_valid()`, and 
`PM_pta()` all have a `$save()` method. This method saves the object to the hard drive
in the current working directory by default. The format is .rds which is a 
binary format used by R to save individual objects. The purpose of the `$save()`
method is to enable retrieval of the object at a later time.


### Loading functions

After a successful NPAG or NPOD run, `PM_load()` creates a 
`PM_result()` object rather than loading run results into the current 
environment and suffixed with the run number as for legacy mode. 

```{r echo=T, eval=FALSE}
res1 <- PM_load(1)
res1$op$plot()
```

For `PM_result()`, `PM_fit()`, `PM_sim()`, `PM_valid()`, and 
`PM_pta()` use the `$new()` method and provide the
full or relative path (if not
in the working directory) and name of an .rds file created by the corresponding
`$save()` method.

```{r echo=T, eval=FALSE}
sim1 <- PM_sim$new("sim.rds")
```



### Report generation

The `PMreport()` function is automatically run at the end of a successful run, 
and it will generate an HTML page with summaries of
the run, as well as the .Rdata files and other objects. The default
browser will be automatically launched for viewing of the HTML report
page. 

## Other functions

Within Pmetrics there are also functions to manipulate data .csv files
and process and plot extracted data.

### Data manipulation

Comparison between the current and the Legacy methods are shown for education.
```{r echo=F}
data_comp <- read.csv("Data/RLcomp_data.csv", na.strings = ".")
knitr::kable(data_comp)
```

### Model selection and diagnostics

Comparison between the current and the Legacy methods are shown for education.
```{r echo=F}
mod_comp <- read.csv("Data/RLcomp_valid.csv", na.strings = ".")
knitr::kable(mod_comp)
```

### Other functions

Comparison between the current and the Legacy methods are shown for education.
```{r echo=F}
oth_comp <- read.csv("Data/RLcomp_other.csv", na.strings = ".")
knitr::kable(oth_comp)
```


<!-- * Process data: `makeAUC`, `makeCov`, `makeCycle`, `makeFinal`, `makeOP`, `makePTA`, -->
<!-- `makeErrorPoly` -->

<!-- * Plot data: `plot.PMcov`, `plot.PMcycle`, `plot.PMfinal`, `plot.PMmatrix`, -->
<!-- `plot.PMop`,`plot.PMsim`, `plot.PMvalid`, `plot.PMpta` -->


<!-- * Pmetrics function defaults: `setPMoptions`, `getPMoptions` -->

Again, all functions have extensive help files and examples which can be
examined in R by using the `help(command)` or `?command` syntax.
