\input amstex
\NoBlackBoxes 
\nopagenumbers

\centerline{{\bf Quickstart Instructions for DVODE.F90}}
\vskip 0.50in

\item{ $\bullet$ } {\bf Compiling DVODE.F90}
\vskip .10cm
\itemitem{ 1. }
DVODE.F90 is a Fortran 90 extension of the f77 DVODE.F
ODE solver of Brown, Byrne, and Hindmarsh from the
LLNL ODEPACK suite.
It incoporates additional options from other solvers
in ODEPACK most notably ones from LSODES and LSODAR.
It also incoporates a Fortran 90 translation of the
Harwell MA28 routines for sparse Jacobians.
The three Fortran 90 modules that constitute DVODE.F90 are
contained in the distribution file vode\_double\_m.f90.
Compile this file using commands appropriate for your
Fortran 90 compiler.
\vskip .05cm
\itemitem{ 2. }
Compile, link, and execute the program example1.f90. The last
line in the output file example1.dat should be:
``No errors occurred.''
Compare your results in example1.dat with those in
Save\_example1.dat.
\vskip .05cm
\itemitem{ 3. }
Repeat Step 2 for the program example2.f90.
Compare your results in example2.dat with those in
Save\_example2.dat.
\vskip .05cm
\itemitem{ 4. }
Refer to the demonstration programs available from the DVODE.F90
support page for examples which illustrate the various features
and options available in DVODE.F90.
If your compiler complains about argument and array declarations
in the user subroutines, refer to example1.f90 and example2.f90
to see how to declare the INTENT of the subroutine arguments.
\vskip .025in
 
\item{ $\bullet$ } {\bf Documentation}
\vskip .10cm
\itemitem{ 1. }
The distribution file vode\_double\_m.f90 contains a detailed
documentation prologue for DVODE.F90.
\vskip .05cm
\itemitem{ 2. }
The DVODE.F90 support page contains a break down of the
documentation prologue by major task.
\vskip .025in

\item{ $\bullet$ } {\bf Changing Working Precision}
\vskip .10cm
\itemitem{ 1. }
The distribution file contains a double precision version of
DVODE.F90. If you wish to use it in single precision change
the selected precision in a statement at the very beginning
of the file.
\vskip .025in

\item{ $\bullet$ } {\bf Web Support Page}
\vskip .10cm
\itemitem{ 1. }
The source code, demonstration programs, and other
material for DVODE.F90 is located at:
\vskip .05cm
\itemitem{ }
\hskip .50in http://www.radford.edu/$\sim$thompson/vodef90web.html
\vskip .025in

\item{ $\bullet$ } {\bf Contact Information}
\vskip .10cm
\itemitem{ 1. }
Please contact the authors if you have any questions
concerning DVODE.F90:
\vskip .01cm
\hskip .50in G.D. Byrne
\vskip .01cm
\hskip .50in gdbyrne847\@yahoo.com
\vskip .05cm
\hskip .50in S. Thompson
\vskip .01cm
\hskip .50in thompson\@radford.edu
\vskip .025in
\itemitem{ 2. }
Note that Alan Hindmarsh is not involved with the distribution
or support of DVODE.F90. {\it Please} do not bug him with
questions about DVODE.F90.

\vfil

\bye
